# 5 May 2017

```{r setup, echo=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
#######################################################################################

# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
library(ProjectTemplate)
load.project()
```

## Calculate Ler spread rates
First we look at furthest distance reached.
```{r furthest}
Ler_furthest_C <- filter(popLer, Treatment == "C") %>%
  group_by(Gap, Rep, Generation) %>%
  summarise(Furthest = max(Pot))
```

Now calculate per-generation spread rates
```{r spread}
# The "default" argument to lag() replaces the leading NA with the specified value
gen_spread <- group_by(Ler_furthest_C, Gap, Rep) %>%
  mutate(speed =  Furthest - lag(Furthest, default = 0),
         speed_m1 = lag(speed))
```

Hmm, there is one case (rep 8, 3p gap) where the forward pot seems to go extinct, resulting in a negative speed. Need to check the original data, but for now let's just set that to zero.

```{r spread_fix}
gen_spread <- within(gen_spread, speed[speed < 0] <- 0)
gen_spread <- within(gen_spread, speed_m1[speed_m1 < 0] <- 0)
```

Look at autocorrelation:
```{r AR1, warning=FALSE}
spread_AR <- group_by(gen_spread, Gap, Rep) %>% 
  summarise(AR1 = cor(speed, speed_m1, use = "complete"))
ggplot(spread_AR, aes(x = AR1)) + geom_histogram(aes(fill = Gap))
```
ALthough it looks like a bias towards negative values, I think that's mostly from the 3p gaps, where the pattern 0-4-0 will be common.

Look at trends by rep:
```{r speed_trends, message=FALSE}
ggplot(gen_spread, aes(x = Generation, y = speed, color = as.factor(Rep))) +
  geom_point(position = "jitter") + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~Gap)
ggplot(gen_spread, aes(x = Generation, y = speed)) +
  geom_point(position = "jitter") + 
  geom_smooth() + 
  facet_wrap(~Gap)
ggplot(gen_spread, aes(x = Generation, y = speed)) +
  geom_point(position = "jitter") + 
  geom_smooth(method = "gam", method.args = list(k = 4)) + 
  facet_wrap(~Gap)
summary(lm(speed ~ Generation, data = filter(gen_spread, Gap == "0p")))
library(mgcv)
summary(gam(speed ~ s(Generation, k = 4), data = filter(gen_spread, Gap == "0p")))
plot(gam(speed ~ s(Generation, k = 4), data = filter(gen_spread, Gap == "0p")))
anova(gam(speed ~ s(Generation, k = 4), data = filter(gen_spread, Gap == "0p")), 
      gam(speed ~ Generation, data = filter(gen_spread, Gap == "0p")),
      test = "Chisq") 
```

So there looks to be some nonstationarity in the continuous runways, with the speed peaking in generation 3. The initial increase makes sense, but the decline does not.
