# 22 September 2017
```{r setup, echo=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
#######################################################################################
# Note that the cache = FALSE is needed to prevent ProjectTemplate errors in the next 
#   chunk. This is unfortunate as it slows build of the full notebook.

# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
library(ProjectTemplate)
load.project()
```

The next steps in the Ler dispersal analysis are to confirm statistical support for among-rep heterogeneity and see if the kernel parameters are correlated with the fraction dispersing.

## Housekeeping
Added the functions for the truncated distributions and to fit the models to lib/helpers.R.

## Kernel heterogeneity
The total AIC for the rep-specific fits was 58230.28 and 58366.17 for the normal and lognormal models, respectively.
