# 26 April 2017

```{r setup, echo=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
#######################################################################################

# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
library(ProjectTemplate)
load.project()
```

## Calculate Ler spread rates
First we look at furthest distance reached.
```{r furthest}
Ler_furthest <- group_by(popLer, Treatment, Gap, Rep, Generation) %>%
  summarise(Furthest = max(Pot))
ggplot(Ler_furthest, aes(x = Generation, y = Furthest, color = as.factor(Rep))) +
  geom_line() + facet_grid(Treatment ~ Gap)
```

Hmm, I've uncovered that popLer has a line of all NA's (and popLer_cm has three)! I've gone ahead and updated `popLer.R` (and added `plyr` to the set of autoloaded packages) to fix this. Now I don't have extra rows/columns in the facet table!

Now let's look at a threshold:
```{r threshold}
threshold <- 100
Ler_thresh <- filter(popLer, Seedlings > threshold) %>%
  group_by(Treatment, Gap, Rep, Generation) %>%
  summarise(Furthest = max(Pot))
ggplot(Ler_thresh, aes(x = Generation, y = Furthest, color = as.factor(Rep))) +
  geom_line() + facet_grid(Treatment ~ Gap)
```

Not surprisingly, this works best for for Treatment C. 

```{r mean_var}
mean(filter(Ler_furthest, Treatment == "C", Gap == "0p", Generation == 6)$Furthest)
var(filter(Ler_furthest, Treatment == "C", Gap == "0p", Generation == 6)$Furthest)
mean(filter(Ler_thresh, Treatment == "C", Gap == "0p", Generation == 6)$Furthest)
var(filter(Ler_thresh, Treatment == "C", Gap == "0p", Generation == 6)$Furthest)
```

The mean and variance of total distance travelled is slightly less in the threshold case, but that might all be attributed to generation 1 (effectively, there is a one-generation delay between inititial colonization and reaching the threshold, here set to 100 seetlings).

HMMPF Now I'm getting an error from load.project() that seems to derive from the 01-Ler-seed-gen1.R --- except that the script works fine when I run it in the console!

4/28: Solved this---it had to do with loading both plyr and dplyr in the initialization file. I've taken out plyr, and explicitly used `plyr::ddply()` in the script that generates popLer.
