# 2 June 2017
```{r setup, echo=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
#######################################################################################
# Note that the cache = FALSE is needed to prevent ProjectTemplate errors in the next 
#   chunk. This is unfortunate as it slows build of the full notebook.

# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
library(ProjectTemplate)
load.project()
```

## Patterns of silique variation, continued
To get around the confusing results last time, let's just create a new factor variable that is Gen:ID, so that each combo is treated independently.

```{r silique_aov}
sil_data <- subset(popLer_sil, Treatment == "B" & Generation > 0 & Gap == "0p")
sil_data$GenID <- with(sil_data, interaction(Gen, ID))
summary(aov(Siliques ~ Gen + GenID, data = sil_data)) 
```
OK, that's better clarity. After accounting for inter-generational variation, most of the remaining variation is *among* runways rather than within runways.

```{r silique_plots, warning=FALSE}
ggplot(aes(Gen, Siliques, group = ID, color = ID), data = sil_data) + geom_point() +
  geom_smooth(aes(Generation, Siliques), span = 1, se = FALSE)
```

Let's look at how the residual variance scales with the mean.
```{r residuals}
sil_stats <- group_by(sil_data, Gen, ID) %>%
  summarise(Mean = mean(Siliques),
            Var = var(Siliques),
            SD = sqrt(Var))
sil_stats <- sil_stats[complete.cases(sil_stats), ]
ggplot(aes(Mean, Var, color = ID), data = sil_stats) + geom_point() 
ggplot(aes(Mean, SD, color = ID), data = sil_stats) + geom_point() 
ggplot(aes(Mean, Var), data = sil_stats) + geom_point() + geom_smooth(method = "lm")
ggplot(aes(Mean, SD), data = sil_stats) + geom_point() + geom_smooth(method = "lm")
```

The relationship betwen the SD and mean looks a bit more linear, although there's still a lot of scatter.

Now let's see if the seed:silique ratio is structured in any way, or is independent among pots.
```{r join}
sil_data <- subset(popLer_sil, Treatment == "B" & Gap == "3p")
sil_data$Generation <- sil_data$Generation + 1
sil_data$Gen <- as.factor(sil_data$Generation)
popLer$ID <- as.factor(popLer$ID)
sil_seed_data <- left_join(sil_data, popLer)
ggplot(aes(Siliques, Seedlings, color = Gen), data = sil_seed_data) + 
  geom_point() + geom_smooth(method = "lm")
```

Within each generation, the number of seedlings in the home pot is independent of the number of siliques!!!

Now include the dispersing seeds in gen 1. 
```{r join_disp}
sil_data <- subset(popLer_sil, Treatment == "B" & Gap == "0p" & Gen == "0")
sil_data$Generation <- sil_data$Generation + 1
sil_data$Gen <- as.factor(sil_data$Generation)
Ler_seed_gen1$ID <- as.factor(Ler_seed_gen1$ID)
sil_seed_data <- left_join(sil_data, Ler_seed_gen1)
ggplot(aes(Siliques, eff_sd_no), data = sil_seed_data) + geom_point() + 
  geom_smooth(method = "lm")
ggplot(aes(Siliques, home), data = sil_seed_data) + geom_point() + 
  geom_smooth(method = "lm")
ggplot(aes(Siliques, stay_prop), data = sil_seed_data) + geom_point() + 
  geom_smooth(method = "lm")
```

So the lack of relationship between siliques and seedling is not because of variability in dispersal.

Let's look at population structure in the variability of home pot seedlings, using the 1p gaps to get replication within populations but seedling numbers mostly from local production:

```{r seed_aov}
seed_data <- subset(popLer, Treatment == "B" & Generation > 1 & Gap == "1p")
seed_data$GenID <- with(seed_data, interaction(Gen, ID))
summary(aov(Seedlings ~ Gen + GenID, data = seed_data)) 
ggplot(aes(Gen, Seedlings, group = ID, color = ID), data = seed_data) + geom_point() +
  geom_smooth(span = 1, se = FALSE)
```

Let's look at how the residual variance scales with the mean.
```{r residuals2}
sil_stats <- group_by(seed_data, Gen, ID) %>%
  summarise(Mean = mean(Seedlings),
            Var = var(Seedlings),
            SD = sqrt(Var))
sil_stats <- sil_stats[complete.cases(sil_stats), ]
ggplot(aes(Mean, Var, color = ID), data = sil_stats) + geom_point() 
ggplot(aes(Mean, SD, color = ID), data = sil_stats) + geom_point() 
ggplot(aes(Mean, Var), data = sil_stats) + geom_point() + geom_smooth(method = "lm")
ggplot(aes(Mean, SD), data = sil_stats) + geom_point() + geom_smooth(method = "lm")
```
So, despite the lack of relationship between siliques and seedlings, we do see that seedling production is structured by runway, and that the SD scales roughly linearly with the mean (althoug not in a way that bives a constant CV)

We can also look at population-level structure in the density-dependent populations.
```{r seed_lm}
seed_data <- group_by(popLer, ID, Pot) %>%
  mutate(Nm1 = lag(Seedlings))
#seed_data$Nm1 <- lag(popLer$Seedlings)
seed_data <- subset(seed_data, Treatment == "C" & Generation > 1 & Gap == "1p")
seed_data$GenID <- with(seed_data, interaction(Gen, ID))
anova(lm(log(Seedlings/Nm1) ~ log(Nm1) + Gen + GenID, data = seed_data)) 
ggplot(aes(log(Nm1), log(Seedlings/Nm1), group = GenID, color = Gen), data = seed_data) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
```
So the population-level variation is still here.
