# (PART) 11 April 2017 {-}

# Getting data in via ProjectTemplate

First, here is a draft of the R header chunk that I will want to include in every post that uses R (silently):

```{r setup}
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
I'm sure I'll come up with some more.

Now, on to how to load the Arabidopsis data using ProjectTemplate. I started with the Ler population data, which is in multiple xlsx files, with one sheet having data, one having metadata, and potentially some blank sheets as well. If I follow the standard protocol for ProjectTemplate and drop them into the `data` directory, then when I execute `load.project()` I get a data object for every sheet of every file (because there is no way to pass options to `read.xlsx()`. In addition, now that they are loaded as data objects, it is hard to access them in a loop to build a the combined datasets. Finally, the workspace is cluttered with all of these objects; if I delete them as part of the data munging, then they will get reloaded the next time I use `load.project()`, which is time consuming!

So instead I took the data files out of the project directory, and used the option to have a `.R` file in the `data` directory. Using this, I can re-use my existing code to read the files in from the Dropbox directory and manipulate them as I want, deleting the intermediate objects and just keeping the final objects. Furthermore, if I then cache the objects, then subsequent loads are fast!

So here's the file:
```{r popLer.R}
cat(readLines('data/popLer.R'), sep = '\n')
```


Now let's look at what we get:
```{r viewenv}
ProjectTemplate::load.project()
ls()
head(popLer)
head(popLer_cm)
```

Looking good!

This has the added advantage that any changes in the data that Jenn makes can be easily loaded by deleting the cache; and that I can share the project on a public github directory (and Jenn can use it as long as we have our Dropbox in the same place).
