# 21 April 2017

```{r setup, echo=FALSE, cache=FALSE}
### This chunk can be deleted from any entry that lacks R code

# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
```{r setup2, echo=FALSE, message=FALSE, cache=FALSE}
### This chunk can be deleted from any entry that lacks R code

# Load the project. For some reason it has to be in a separate chunck.
library(ProjectTemplate)
load.project()
```

## Data munging
Realizing that I need to re-use the `seedlings1` data from last time, I see that I should start to use the `munge` directory for key derived data. I therefore ned to give it a more meaningful name; how about `Ler_seeds_gen1`. I'll also add some more information about home and away

The code for doing this is in `munge/01-Ler-seed-gen1.R`

## Ler fecundity
In the process of doing this I discovered that last time I failed to double the dispersing seeds in the effective seed number calculation. So let's redo the graph and table.

Plot the distributions:
```{r sdlgHist1}
ggplot(Ler_seed_gen1, aes(x = eff_sd_no, fill = as.factor(Density))) + 
  geom_histogram(binwidth = 50)
```

(of course, stacked histograms aren't great, but it gives the general idea).

Now look at the mean and variance across pots:
```{r sdlgTable1}
kable(group_by(Ler_seed_gen1, Density) %>% 
        summarise(Mean = mean(eff_sd_no), Variance = var(eff_sd_no)),
      caption = paste("Mean and variance across pots of effective seed number",
                      "in treatments B and C of Ler generation 1")
      )
```

This changes the quantitative mean:variance ratios:

- For isolated plants, the variance is about `r round(3443/134)` times the mean. Compared with Poisson, this is huge!
- As we go from a density of 1 to 50, the seeds per plant drops by a factor of `r round(134/(348/50))`, from 134 to about `r round(348/50)`.
- If the among-pot variance at high density was caused by iid variation in individual seed production, the among individual variance would need to be 21363/50, or about `r round(21363/50)`. 
    - This is a `r round((21363/348) / (3443/134))`-fold increase in the variance:mean ratio
    - It seems that, to acheive this among-individual variance, almost all the seeds in the high density plots would need to come from one or two individuals--and whether the among-successful-individual variance required is consistent with what we see in isolated pots remains an open question. We would need to assume something like a lognormal distribution among individuals to get at this.

`r var(c(134, rep(0, 49)))`
