# 31 May 2027
```{r setup, echo=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
#######################################################################################
# Note that the cache = FALSE is needed to prevent ProjectTemplate errors in the next 
#   chunk. This is unfortunate as it slows build of the full notebook.

# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = "..")
```
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
library(ProjectTemplate)
load.project()
```

## Silique counts in the Ler population data
I've loaded in the silique count data from ~Arabidopsis/analysis/*_siliques.csv. I'm not actually sure how "clean" these are; I had to add the gap treatments to generation zero, and in the process found one error in generation 1; I didn't look at the other generations.

The script is in `data/popLer_sil.R` and the data frame is `popLer_sil`. 

## Patterns of silique variation
Let's look at silique production in treatment B as a function of replicate (here labeled "ID") and generation. If my speculations last week hold, then ID:Gen should be significant, but probably not the main effect of ID.

In order to have enough replication to get at the interaction, I'll look only at the continuous runways and not at generation zero.

```{r silique_aov}
sil_data <- subset(popLer_sil, Treatment == "B" & Generation > 0 & Gap == "0p")
summary(aov(Siliques ~ Gen * ID, data = sil_data)) 
```
So I'm not sure how to interpret the main effect of ID. Let's look at the coefficients of a regression to see.
```{r silique_lm}
summary(lm(Siliques ~ Gen * ID, data = sil_data)) 
anova(lm(Siliques ~ Gen * ID, data = sil_data))
```
Oh, I see - aov still uses the default contrasts.
```{r silique_aov2}
## Set orthogonal contrasts.
op <- options(contrasts = c("contr.helmert", "contr.poly"))
summary(aov(Siliques ~ Gen * ID, data = sil_data)) 
options(op)  # reset to previous
```
Nope, that didn't do it either.

