---
title: "Parameter estimation"
author: "Bruce E. Kendall"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
toc: TRUE
bibliography: ../Bibliography.bib
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Knitr options. Note that these mean:
#   - All output will be cached. This is desireable in case code/data changes in the 
#     future; but it means that all chunks should be labeled for reliable behavior. 
#     DEPENDENCIES ACROSS CODE CHUNKS MAY BE AN ISSUE.
#   - R code will run in the project root directory
project_root <- normalizePath("../..")
knitr::opts_chunk$set(comment = '', cache = TRUE)
knitr::opts_knit$set(root.dir = project_root)


# Clean up
rm(project_root)
```
```{r loadproject, echo=FALSE, message=FALSE}
library(ProjectTemplate) 
load.project()
```

This document describes the parameterization of the models in Kendall, Williams and Levine. 
There are two sets of parameters: a single genotype model, parameterized with Ler [@Williams2018]; and a multi-genotype model, parameterized with the "RIL" data [@Williams2016b].
For now we include all code, so that we can see what's going on; if we need to share this as a supplement then we will need to clean it up.

# Single-genotype model
## Dispersal kernel
Our dispersal kernel is essentially zero-inflated: a certain fraction of seeds stay in the maternal pot, and the remainder disperse according to a standard kernel. We only account for seeds staying in the runway. 

We used a beta-binomial model to estimate the fraction of seeds dispersing from the maternal pot: the binomial part accounts for independent dispersal by individual seeds, and the beta part accounts for heterogeneity among pots in the expected fraction dispersing.

We modeled the distance travelled by dispersing seeds using a number of continuous distributions:

* Log-normal
* Half-normal
* Exponential
* Gamma
* Wiebull
* Inverse Gaussian
* Logistic
* Inverse gamma
* Generalized gamma

### Objectives
1. Estimate mean and variance of fraction dispersing, based on beta-binomial model
2. Determine which dispersal kernel is "best"
3. Estimate kernel parameters for each replicate
4. Examine whether dispersal parameters depend on density
5. Estimate mean, variance, and distributional form for kernel parameters

### Data source
Data on dispersal come from a dedicated seed dispersal experiment, which is described by @Williams2018:
> To quantify dispersal, as well as the potential effect of density on dispersal, we used data from a second experiment, where we grew plants across a range of densities (N = 48 pots, containing between 1 and 500 individuals) under the same growing conditions as in the main experiment. During the seed dispersal phase, a sheet of sticky paper as wide as the pots (7.3 cm) was placed to one side of each pot at the same height as the pot, and all seeds that landed on the sticky paper were counted, as well as all seeds that germinated in the starting pot. (p. 878)

The primary data object for this analysis is called `disperseLer`, which is derived from `2013_08_08_Exp1_Spray.csv`:
```{r dispersler}
cat(readr::read_file("data/disperseLer.R"))
summary(disperseLer)
head(disperseLer)
```
The `Seedlings` column records the number of seedlings that germinated in the home pot when associated with `Distance` value of 4; at other distance values it is the number of seeds found in the one-cm strip from `Distance` - 1 to `Distance`.
Thus, `Distance` = 8 is the first strip on the sticky paper, and represents dispersal 0--1 cm beyond the maternal pot.

### Methods
There was a great deal of heterogeneity among pots in the fraction dispersing, more than would be expected from binomial variability from a finite sample of seeds.
Therefore, we fit a beta-binomial model to the data.
For statistical robustness, we used `vglm()` with a `betabinomial` distribution from the **VGAM** package [v. `r packageVersion("VGAM")`; @Yee2015], fitting an intercept-only model.
The parameters of the fitted beta distribution are reported as $\mu$, the mean of the distribution, and $\rho$, a shape parameter.
The variance of the distribution was calculated as $V = \mu (1-\mu) \rho$.
The analysis code is as follows:
```{r betabinom}
cat(readr::read_lines("munge/10-make_Ler_params.R", skip = 4, n_max = 19), sep="\n")
```

The values in the data frame implicitly assign distance zero to the trailing edge of the pot; other plausible candidates are the middle of the pot (hence the first band in the dispersal data is from 3.4 to 4.5) or the leading edge of the pot (hence the first band in the dispersal data is from 0 to 1).
For the kernel estimation, it is not obvious what the "correct" value should be, as the plants that are the source of the seeds are spread throughout the 7 cm wide pot (but note that they might not all contribute equivalently to dispersing seeds: plants on the edge of the pot often "lean" over the disperal area if it is not already occupied by plants, and dispersal from plants in the middle of the pot may be impeded by the edge plants.
From a practical standpoint, we test all three, and choose the one that seems to provide the best general fit. 
IN PRACTICE, I SET DISTANCE ZERO TO THE LEADING EDGE OF THE POT, AND DID NOT DO ANY TRUNCATION AS DESCRIBED BELOW.

The dispersal data are "interval censored"---that is, we don't know where in the one-cm bands (or the maternal pot) the seeds lie. 
We fit distributions using the **fitdistrplus** package [v. `r packageVersion("fitdistrplus")`; @Delignette-Muller2015], which can account for interval and terminal truncation of the data.

I HAVE NOT ACTUALLY DONE THE FOLLOWING, AND AM NOT SURE I WANT TO.
Furthermore, when we set the origin to values other than the leading edge of the pot, then we essentially have not observations from the left end of the distribution (since we don't know what fraction of the seeds in the maternal pot are "didn't disperse" and what fraction are "dispersed less than one pot's distance").
In those cases we need to truncate the distributions so that the probabilities are calculated correctly.
This also allows us, in simulations, to not have to distinguish two classes of non-dispersing seeds.
The fitting routine we use requires p and d forms for the density and distribution functions of the distribution.
Here are examples for a left-truncated lognormal:
```{r tlnorm}
dtlnorm
ptlnorm
```

We fit all of the models to each of the replicates, and examined the values of $\Delta$-AIC to determine if there was a distribution that consistently fit best, or at least well.
These models are fit using the `fit_dispersal_untruncated` function:
```{r}
source("lib/helpers.R")
fit_dispersal_untruncated
```
It makes use of the following helper functions:
```{r}
cens_dispersal_data
start_params
start_gengamma
```

### Fraction dispersing
To estimate the mean and standard deviation of the beta-binomial model, we source the code listed above:
```{r}
frac_disp_script <- paste(readr::read_lines("munge/10-make_Ler_params.R", 
                                            skip = 4, n_max = 19), 
                          collapse = "\n")
source(textConnection(frac_disp_script))
```
We can look at the full output of the vglm:
```{r}
summary(bbfit)
```
$\mu$ and $\rho$ are `r VGAM::Coef(bbfit)[1]` and `r VGAM::Coef(bbfit)[2]`, respectively.
After transformation, the mean fraction dispersing is `r frac_dispersing` and the standard deviation in the fraction dispersing is `r fd_stdev`.

### Comparing dispersal kernels
For each of the dispersal replicates, fit each dispersal kernel and record the AICs. This is done assuming that "distance zero" is at the leading edge of the maternal pot.
```{r fitallkernel, warning=FALSE, error=FALSE, message=FALSE}
result <- fiteach_disp_unt(disperseLer) 
```
Here are the fits for the first few replicates:
```{r, message=FALSE}
library(tidyverse)
result <- group_by(result, ID) %>% mutate(delta_AIC = AIC - min(AIC))
print(result, n=27) 
```

Here's a plot of the $\Delta$-AIC values for each model:
```{r, message=FALSE}
ggplot(result, aes(x=delta_AIC, group = model)) + 
  geom_histogram() + 
  facet_wrap(~model, scales = "free") 
```

Notive that the $\Delta$-AIC values for the generalized gamma ("gengamma") distribution are all less than 3.5, whereas all the other distributions have a substantial number of values larger than 10, and often very large indeed.
Thus, we select the generalized gamma distribution as the best distribution to use.

There are two replicates that the generalized gamma distribution refuses to fit.
These are 79_0 and 90_1.
Let's take a look at them:
```{r badfit}
filter(disperseLer, ID =="79_0")
filter(disperseLer, ID == "90_1")
```
The first problematic replicate only has 5 dispersing seeds, so it's not surprising that it doesn't find a good fit.
It's unclear what the problem is with the second one, but continued fiddling has not been able to isolate and fix the problem.
So for the remainder of the analysis we drop those two replicates.
```{r dropbadfit}
disperseLer2 <- filter(disperseLer, ID != "79_0", ID != "90_1")
```

### Replicate-specific kernel parameters
Here are the replicate-specific fits:
```{r fiteach}
fiteach <- fiteach_disp_unt(disperseLer2, model = "gengamma")
fiteach
```

The shapes of the kernels look quite different:
```{r ploteach, warning=FALSE}
p <- ggplot(data = data.frame(x = c(0, 20)), mapping = aes(x = x)) 
for (i in 1:nrow(fiteach)) {
  pvec <- as.numeric(fiteach[i, 4:6])
  plist <- list(mu = pvec[1], sigma = pvec[2], Q= pvec[3])
  p <- p + stat_function(fun = dgengamma, args = plist)
}
p + xlab("Distance (cm)") + ylab("Probability density")
```

And here is a scatterplot of the parameters:
```{r fiteach_scatterplot}
params <- fiteach[, 4:6]
names(params) <- c("mu", "sigma", "Q")
pairs(params)
```

Notice that one of the estimates looks like an outlier.
It is replicate 131_0.
Let's look at it:
```{r wierdfit}
filter(disperseLer, ID =="131_0")
```
There does not appear to be anything particularly odd about it, but it will have an outsized influence on the covariance between parameters, so we drop it.
```{r fiteach2}
disperseLer2 <- filter(disperseLer2, ID != "131_0")
fiteach <- fiteach_disp_unt(disperseLer2, model = "gengamma")
```

To check for heterogeneity, compare the summed AIC of the individual fits to the AIC of a single fits to combined data.
Here is the combined fit:
```{r fitall}
fitall <- fit_dispersal_untruncated(disperseLer2, model = "gengamma")
fitall[, -(1:2)]
```
The summed AIC of the rep-specific fits is `r round(sum(fiteach$AIC))`, which is lower than the homogeneous fit by `r round(fitall$AIC[1] - sum(fiteach$AIC))` AIC units.
Strong evidence for heterogeneity!

### Density dependence in dispersal
The maternal pots for this experiment had different densities of plants; this is recorded in the `Density` column of `disperseLer`.
We can examine density-dependence in the fraction dispersing by including Density as an explanatory variable in the call to `vglm()`:
```{r frac_dispersing_dd}
nondispersers <- subset(disperseLer, Pot == 0, c("ID", "Seedlings", "Density"))
dispersers <- filter(disperseLer, Pot >= 1) %>%
  group_by(ID) %>%
  summarise(dispersers = sum(Seedlings))
disperse_num <- merge(nondispersers, dispersers)
names(disperse_num)[2] <- "nondispersers"
bbfit_dd <- vglm(cbind(dispersers, nondispersers) ~ Density,
              betabinomial,
              data = disperse_num)
summary(bbfit_dd)
```
The coefficient on Density is clearly not different from zero.

For the dispersal kernels, we can get a preliminary sense by looking at the relationship between density and each of the three parameters.
```{r kernel_dd}
fiteach_dens <- merge(fiteach, disperse_num)
ggplot(fiteach_dens, aes(x = Density, y = par1)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  ylab("mu")
ggplot(fiteach_dens, aes(x = Density, y = par2)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  ylab("sigma")
ggplot(fiteach_dens, aes(x = Density, y = par3)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  ylab("Q")
```

It's pretty clear that there's nothing going on here, so no need for fancy modeling in the original fits.

### Joint kernel parameters
We will model the among-rep variability as multivariate normal (MVN), with the means of the three parameters and the covariances omong them.
Add better names:
```{r rename}
names(fiteach)[4:6] <- c("mu", "sigma", "Q")
```
Calculate the mean and covariance matrix:
```{r cov}
apply(fiteach[, 4:6], 2, mean)
cov(fiteach[, 4:6])
```

## Seed production
For seed production, we use the Gompertz model:
$$\log \frac{N_{t+1}}{N_t} = a + b \log N_t.$$
Now, in principle we could estimate these parameters by doing a GLM of $N_{t+1}$ on $\log N_t$ using a log link; subtracting one from the slope gives an estimate of $b$.
However, preliminary exploration of tte data revealed three wrinkles [see plots and analysis from 5 June 2017, which could be added here]:

1. There is temporal structure in the residuals, with the residual variance depending on generation. 
This represents temporal variability in expected seed production (often called "environmental stochasticity") that needs to be both accounted for and estimated for the model.
2. Even after accounting for generation, there is structure in the residuals related to replicate: the pots within a replicate are more similar to each other than they are to pots in other replicates. 
This is a form of spatial variability that also needs to be estimated.
3. After accounting for the temporal and spatial structure, the remaining residual variation among pots (which in the model we think of as representing demographic stochasticity) is hyperdispersed relative to Poisson residuals.

The first two issues are addressed using random effects.
The third could be addressed using a quasi-Poisson link function, but that doesn't play well with `glmer()`, and is not a distribution we can use in the simulation, as there is no `rquasi` function.
Instead, we calculate a variance inflation factor, $\theta$, which can be used with the negative binomial distribution to approximate the quasi-Poisson.

### Data source
This is the one place where we use data from the experimental populations to estimate parameters, as otherwise there would be no possibility of residual variation among pots depended on time or on location in the greenhouse.
The population data are described by @Williams2018 (p. 878):

> Plants were given the opportunity to spread across a linear series of soil-filled plastic planting pots that were 21 $\times$ 7.3 $\times$ 5.5 cm (Bachmann Plantec, Hochdorf, Switzerland), arranged so that each patch was 7.3 cm wide. The total length of each array was extended as the invasions moved, but always extended 58.4 cm (8 pots) past the last pot with reproducing individuals. To assess the influence of landscape patchiness on spread, we separated pots in the arrays by distances of 0 (continuous), 2, 4, or 6 $\times$ mean dispersal distance. Seeds that fell into gaps between patches died. Given the constant patch (pot) size in all patchiness treatments, when we altered the distance between suitable habitats, we also altered the total fraction of suitable habitat.

> We evaluated the influence of density dependence in our experiment by creating the extreme case where density dependence is so strong that the population in a given pot, no matter how dense, can only produce a seed number equivalent to that produced by one individual, which in our experiment is equivalent to a carrying capacity of roughly 43 individuals. Our manipulation was equivalent to setting $\alpha = 1$ in the common annual plant population growth function with Beverton–Holt density dependence\dots. We achieved $\alpha = 1$ in our experiment by thinning all germinating seedlings in each pot down to one solitary individual, so that the number of seeds produced by each patch in this treatment was equivalent to the number of seeds produced by one individual. We refer to this treatment as "amplified density dependence" throughout the paper\dots. Where we did not manipulate density dependence ("control"), the empirically measured $\alpha$ was 0.078, or less than one twelfth of the amplified value.

> The two treatments (4 gap sizes and 2 levels of density dependence) were applied in a factorial design to test for their interaction. For each combination, we established 10 replicate arrays, giving a total of 80 replicates, though one replicate was lost during the experiment due to a fallen pot. We started the experiment by sowing 50 seeds in the left-most pot in the unmanipulated density dependence arrays, and 3 seeds in the amplified density dependence arrays (thinned to 1 individual after germination). In subsequent generations in the amplified density dependence treatment, after germination, and after the number of seedlings was counted, we thinned all pots to one individual. Survival was nearly 100%, but if the remaining individual subsequently died, it was replaced with a new plant. When there were multiple individuals in the leading edge pot, we always chose the individual farthest from the origin that was in the center of the pot ($\pm 3$ cm from the midline). We chose the farthest so as to not truncate the distance seeds had dispersed, but we also chose based on centrality to prevent leading individuals growing in corners, where a large fraction of seeds would leave the arrays altogether. In pots behind the leading edge, we chose the central-most individual. In both treatments, germinants grew, made seeds, and seed dispersal was assisted with a simulated rain event that aided in both opening fruits through wetting and drying them and in dislodging seeds from the central partition of each fruit. Specifically, after plants were given 6 weeks to drop their seeds unassisted, we sprayed all patches from directly above with a spray bottle at a height of 60 cm in a systematic way to ensure each patch received exactly four sprays, and plants were then given a further week to disperse seeds. For more details on growing conditions, see @Williams2016b. We ran the experiment for 7 generations, allowing for 6 generations of dispersal.

The population data are in a series of spreadsheets, one for each generation.
First we process the data to get them into a single data frame that is amenable to analysis (this is used in other parts of the project as well).
```{r popLer}
cat(readr::read_file("data/popLer.R"))
```

Then we further process the resulting `popLer` data frame, to do several things:

* Restrict data to treatments B ("augmented density dependence") and C ("control")
* Focus on replicates having one- or two-pot gaps
* Create a lagged seedling abundance variable
* Select replicate $\times$ generation cases with $> 1$ pots (to allow for distinguishing between within-rep and among-rep variation)

```{r Ler_fecundity}
cat(readr::read_file("munge/05-Ler_fecundity.R"))
summary(Ler_fecundity)
head(Ler_fecundity)
```

The logic here is that in the fragmented landscapes, the vast majority of seedlings in pot $i$ in generation $t$ were produced by parents in the same pot the previous generation; the number of seeds dispersing from other pots is small enough and variable enoug that it can be treated as noise.
We do not use the three-pot-gap treatment because very few populations spread beyond the first pot; while these pots could be useful contributions for estimating the mean density-dependent effect, they do not allow for distinguishing between among- and within-runway variation.

Finally, the seedling numbers in this dataset represents the seeds that did not disperse.
Since we don't have information on dispersal from these pots, we augment the seed production by the mean fraction dispersing, calculated above:
```{r augment_seeds}
Ler_fecundity$Seedlings <- round(Ler_fecundity$Seedlings / frac_dispersing)
```

### Parameter estimates

Here's the estimation.
We use the **lme4** package [@Bates2015] to do the mixed effects model.
```{r Ler_seed}
# Fit model
library(lme4)
seeds_DD <- glmer(Seedlings ~ log(Nm1) + (1 | Gen) + (1 | GenID), 
                  data = Ler_fecundity, family = poisson)
summary(seeds_DD)

# Extract parameters
Ler_params <- list(
  a_Gompertz = fixef(seeds_DD)[1],
  b_Gompertz = fixef(seeds_DD)[2] - 1,
  sigma_seed_time = as.data.frame(VarCorr(seeds_DD))$sdcor[2],
  sigma_seed_rep = as.data.frame(VarCorr(seeds_DD))$sdcor[1],
  theta_seed = sum(residuals(seeds_DD, type="pearson")^2)/df.residual(seeds_DD)
)

Ler_params
```

# References {#references}
