---
title: Build site broken
author: Bruce Kendall
date: '2017-12-27'
slug: build-site-broken
categories:
  - Housekeeping
tags:
  - blogdown
---

Yesterday I updated **blogdown**, from v. 0.1 to 0.4. This broke `build.site()`, because when it tries to compile an Rmd file that calls `projecttemplate`, the working directory is no longer being set to the project root, which presently is "/Users/kendall/Documents/Github/InvasionHet". This was acheived by setting the knitr parameter `root.dir` in `.Rprofile` in the post directory. 

Here's what the knitr option currently is:
```{r project.root, cache=FALSE}
knitr::opts_knit$get()$root.dir
# Originally gave NULL
```
and here's what is being viewed as the working directory:
```{r pwd, cache=FALSE}
getwd()
# Originally gave post subdirectory
```
So it appears that the local `.Rprofile` is no longer being read in.

When I move the `.Rprofile` to the project root (and update the relative path to point to that same directory) then it *does* get read in (see updated values above).

Well, that didn't fix the problem. The call to `ProjectTemplate::load.project()` (which is first encountered in `content/post/2017-09-19-fit-ler-dispersal-kernels.Rmd`) still sees the current directory as `post`. In fact, it's not even clear why the chunk was executing, as cache was set to TRUE!

Maybe the cache location has been moved?

What's even wierder is that, having modified 2017-09-19-fit-ler-dispersal-kernels.Rmd (just turning caching on and off), it re-rendered under serve.site just fine!

I don't know why things aren't recompling fine under build_site. But the workaround for creating a publishable version without rebuilding the Rmd files is to use `blogdown::hugo_build()` instead.
