---
title: fitdistrplus
author: Bruce Kendall
date: '2018-10-04'
slug: fitdistrplus
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
#library(ProjectTemplate)
ProjectTemplate::load.project()
#print(parent.frame())
```
I've modified the helper functions to separate out the creation of the censored data frame from the fitting, and to allow setting the origin. Note that setting `zero = 7` only looks at the runway, and allows use of non-truncated distributions when the distribution only has support for non-negative values.

For example:
```{r}
library(fitdistrplus)
temp <- filter(disperseLer, ID == "100_0")
cens_data <- cens_dispersal_data(temp, 7)
p1 <- fitdistcens(cens_data, "lnorm")
summary(p1)
plot(p1)
p2 <- fitdistcens(cens_data, "exp")
summary(p2)
plot(p2)
```

I've just found that the **rmutil** package adds a lot of distributions, some of which may be useful (not 2DT though). However, starting values of the parameters need to be set. Furthermore, there appears to be a bug in `fitdistcens`: comparing with `fitdist`, it appears that the line
```
    arg_startfix <- checkparamlist(arg_startfix$start.arg, arg_startfix$fix.arg, 
        argddistname)
```
needs to be replaced by
```
    hasnodefaultval <- sapply(formals(ddistname), is.name)
    arg_startfix <- checkparamlist(arg_startfix$start.arg, arg_startfix$fix.arg, 
        argddistname, hasnodefaultval)
```

I can hack that, but need to send a bug report (unfortunately the code is on R-forge and not very user-friendly).
