---
title: More on distributions
author: Bruce Kendall
date: '2018-11-20'
slug: more-on-distributions
categories:
  - Code development
  - Parameter estimation
tags:
  - dispersal
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```


```{r data0}
temp <- filter(disperseLer, ID == "100_0")
cens_data <- cens_dispersal_data(temp, 7)
```

## Generalized gamma
I've made a specialized start function for gengamma:
```{r start_gg}
start_gengamma
```
This can called via `start_params()`:
```{r start_gg2}
start_params
```
Note that the optional parameter `truncated` can be passed through via the dots. The truncated case needs to be written (the function in fitdistcens will be different).

Note that there are comments in the code that are not coming through here---somehow the rmarkdown rendering is "tidying" the code output!
