---
title: Likely-looking distributions
author: Bruce Kendall
date: '2018-11-16'
slug: likely-looking-distributions
categories:
  - Conceptual model evaluation
  - Parameter estimation
tags:
  - dispersal
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
#library(ProjectTemplate)
ProjectTemplate::load.project()
#print(parent.frame()) 
```

Let's think about which distributions could be good. We know what the (half-)normal, exponential, and log-normal look like. 

Nathan et al. (2012) also suggest:

* Exponential power
* 2Dt
* Inverse power
* Logistic
* Mixture models
* Inverse Gaussian (Wald)
* Wiebull
* Gamma

The scaling factors in Nathan et al. are for two-dimensional dispersal expressed in radial distance (I think). They suggest that this can be converted to a "distance kernel" by multiplying by $2\pi r$, but that doesn't seem right for e.g. their Gaussian formulation.

Some notes:

* Gamma is generalization of exponential. It can have a mode at the origin or at positive values. Good candidate.
* Inverse Gaussian is also a good candidate, based on shape. 
* Both gamma and inverse Gaussian are special cases of the 3-parameter generalized inverse Gaussian. `GIGrvg::(d)gig` (but no p form); `rmutil::(d)ginvgauss`
* The exponential, Weibull, and gamma are all special cases of the 3-parameter generalized gamma distribution. At least for the plots shown on wikipedia, it does not appear to allow a mode away from zero(?). `flexsurv::dgengamma`

Here is the list of distributions for which start values are found automatically from the help for `mledist`): For the following named distributions, reasonable starting values will be computed if start is omitted (i.e. NULL) : "norm", "lnorm", "exp" and "pois", "cauchy", "gamma", "logis", "nbinom" (parametrized by mu and size), "geom", "beta", "weibull" from the stats package; "invgamma", "llogis", "invweibull", "pareto1", "pareto", "lgamma", "trgamma", "invtrgamma" from the actuar package.

Let's look at some of these, with one of the datasets. We'll start with the origin at the leading edge of the home pot, so I don't have to deal with the left-truncation (I have an idea for generalizing that code).

```{r data0}
temp <- filter(disperseLer, ID == "100_0")
cens_data <- cens_dispersal_data(temp, 7)
```

## Gamma distribution
```{r gamma}
fit_gamma <- fitdistcens(cens_data, "gamma")
summary(fit_gamma)
plot(fit_gamma) 
```

## Logistic distribution
```{r logistic}
fit_logis <- fitdistcens(cens_data, "logis")
summary(fit_logis)
plot(fit_logis) 
```

It looks like this distribution is not restricted to postive values, so needs to be truncated anyway. It seems unlikely to be a good generic fit as it never allows a mode away from zero (just a plateau).

## Weibull distribution
```{r weibull}
fit_weibull <- fitdistcens(cens_data, "weibull")
summary(fit_weibull)
plot(fit_weibull) 
```

## Inverse Gaussian
This requires setting start values to the parameters. Following the logic of other functions in **fitdistrplus**, we'd like to match empirical moments. Wikipedia (https://en.wikipedia.org/wiki/Inverse_Gaussian_distribution) gives a distribution in terms of $\mu$ and $\lambda$, where the mean is $\mu$ and the variance is $\mu^3/\lambda$. The function in **actuar** has parameters "mean" and "shape." It is described in terms of of the "dispersion" parameter which is the inverse of shape; inspection reveals that "mean" is $\mu$ and "dispersion" is $1/\lambda$ in the Wikipedia formulation, so "shape" is $\lambda$. Thus we want to estimate start values using `mean = mean(x)` and `shape = mean(x)^3 / var(x)`.

```{r invgauss}
library(actuar)
x <- apply(cens_data, 1, mean) # set each value to the middle of its interval
start_pars <- list (
                      mean = mean(x),
                      shape = mean(x)^3 / var(x)
                    )
fit_invgauss <- fitdistcens(cens_data, "invgauss", start = start_pars)
summary(fit_invgauss)
plot(fit_invgauss) 
```

I've created a function `start_params()` in `helpers.R` that does this:
```{r start_params, echo=FALSE}
start_params
```
I will add other distributions in the `switch` function. Here's a test that it works:
```{r invgauss2}
library(actuar)
start_pars <- start_params(cens_data, "invgauss")
fit_invgauss <- fitdistcens(cens_data, "invgauss", start = start_pars)
summary(fit_invgauss)
```

## Generalized gamma distribution
This is provided by **flexsurv**, with root `gengamma`. This implementation has parameters $\mu$, $\sigma$, and $Q$.

The Wikipedia definition (https://en.wikipedia.org/wiki/Generalized_gamma_distribution) is 
$$
f(x | a, d, p) = \frac{(p/a^d)x^{d -1}e^{-(x/a)^p}}{\Gamma(d/p)}
$$
The Wikipedia article includes the conversion to the flexsurv definition:
$$
\begin{aligned}
\mu &= \ln a + \frac{\ln d - \ln p}{p}\\
\sigma &= \frac{1}{\sqrt{pd}}\\
Q &= \sqrt{p/d}
\end{aligned}
$$
We actually need to invert these to do the substitutions in the moment calculations. We have
$$
\begin{aligned}
d/p &= Q^{-2}\\
pd &= \sigma^{-2}\\
d/p &= (\sigma p)^{-2} = Q^{-2}\\
p &= Q/\sigma\\
d &= 1/(\sigma Q)\\
\ln a &= \ln (d/p)^{-p} - \mu = \ln Q^{2 Q/\sigma} - \mu\\
a &= \exp\left[ \ln Q^{2 Q/\sigma} - \mu\right]
\end{aligned}
$$
But actually, this isn't going to help much, because the terms for the mean and variance on Wikipedia aren't going to be invertible (at least not easily).

From the description on the help page, it seems like $\mu$ is the mean of $x$ and $\sigma$ is the sd of $\log(x)$. Let's give this a try:
```{r ggtest}
library(flexsurv)
x <- rgengamma(1000, 2, 3, 0)
mean(x)
mean(log(x))
sqrt(var(log(x)))
```
When $Q=0$ (the lognormal limit) then $\mu$ is the mean of $\log(x)$ and $\sigma$ is the sd of $\log(x)$. But the meanings are rather different at the Weibull and gamma limits, as described in the help for gengamma. So I think the strategy will be to fit those three standard parameters and use the AIC-best to set the start values for `gengamma`.
