---
title: Working with fitdistrcens
author: Bruce Kendall
date: '2018-10-18'
slug: working-with-fitdistrplus
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
#library(ProjectTemplate)
ProjectTemplate::load.project()
#print(parent.frame())
```
So here's what happens if we try to run `fitdistrcens` on a non-standard distribution:
```{r fitdistcens, message=TRUE, error=TRUE}
library(rmutil)
temp <- filter(disperseLer, ID == "100_0")
cens_data <- cens_dispersal_data(temp, 7)
try(fitdistcens(cens_data, "ggamma", start = list(s = 5, m = 4, f = 2)), outFile = stdout())
```

Does it work with `fitdist`, which doesn't account for the censoring?
```{r fitdist, message=TRUE, error=TRUE}
p1 <- fitdist(temp$Distance - 7, "ggamma", start = list(s = 5, m = 4, f = 2))
```
No, but gives a different error. I've traced this to the firstline of `computegetparam`, which has first line:
```r
nonparaminR <- c("x", "p", "q", "n", "log", "log.p", "lower.tail", 
        "ncp")
```
However, the first argument of the distribution functions in **rmutil** is "y". I can try to fix that here, but who knows whether the assumption of the first argument being x is baked in anywhere else!

The thing to try is to make a locally modified version of `computegetparam` and, if that works, `fitdistccens`. These will have to be sourced after **fitdistrplus** is loaded.

OK, I've made a modified version of `computegetparam` and put it in `src/myfitdistrplus.R`. Let's try it:
```{r fitdist2, message=TRUE, error=TRUE}
source("src/myfitdistrplus.R")
p1 <- fitdist(temp$Distance - 7, "ggamma", start = list(s = 5, m = 4, f = 2))
```
Shoot, it seems that the local function definition is not overriding the definition within the package. The same thing happens when I do it from the command line, so it's not just a knitr thing.

I think the only solution (short of submitting a bug report and hoping for a fast update) is to download the package from its github mirror (https://github.com/cran/fitdistrplus) and build my own version.
