---
title: rmutil and fitdistrplus
author: Bruce Kendall
date: '2018-10-22'
slug: rmutil-and-fitdistrplus
categories:
  - Parameter estimation
  - Code development
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
#library(ProjectTemplate)
ProjectTemplate::load.project()
#print(parent.frame())
```

So the fix I did last week to `computegetparam()` (to allow "y" as an argument to the distribution functions) does allow `fitdist` to work (sort of) with the **rmutil** distributions. However, there can still be failures in estimatation or SE calculation, because there are a number of other ways that the **rmutil** distributions work differntly from the base distribution functions. **fitdistrplus** expects that:

- If the distribution parameters are out of bounds, `ddist` and `pdist` will return `NaN` (**rmutil** throws an error)
- If `x` is outside the support of `dist`, `ddist` will return zero (**rmutil** throws an error)
- If `x` is less than the minumum supported value of `dist`, `pdist` will return zero (**rmutil** throws an error)
- If `x` is greater than the maxumum supported value of `dist`, `pdist` will return one (**rmutil** throws an error)
- If `x` is NA, `ddist` and `pdist` will return an NA (**rmutil** throws an error)

So I'm going to need to rewrite the distribution functions. Fortunately they are in pure R---see https://github.com/swihart/rmutil/blob/master/R/dist.r. However, the rewriting will be painful enough that I'll want to be selective about the distributions I pick!

Note that I can change y to x; but I still need to make the fix in `fitdistcens`, so will still need my local copy of **fitdistplus**. And because it needs to call functions that are not revealed outside the package environment, it needs to be part of a complete package.

I did just find in the **fitdistr** documentation that, in addition to distributions in the **stats** package, **fitdistr** provides inbuilt support (including initial values) for "invgamma", "llogis", "invweibull", "pareto1", "pareto", "lgamma", "trgamma", "invtrgamma" from the **actuar** package.

I made the change to `fitdistcens`, and it is now getting past the initialization phase. But it still gets hung up because of the errors returned by `dggamma` and pggamma`.

Here are the updated functions:
```{r ggamma}
dggamma
pggamma
```

And here's a fit:
```{r fitdistcens}
temp <- filter(disperseLer, ID == "100_0", Distance > 4)
cens_data <- cens_dispersal_data(temp, 7)
p1 <- fitdistcens(cens_data, "ggamma", start = list(s = 1.7, m = 11, f = 2))
summary(p1)
plot(p1) 
```
The fit looks pretty good, but the correlation matrix suggests that it's overparameterized. Also, not all start values for the parameters work!

Note that, if I want to use an offset different than 7, I'll need to write truncated versions as well.
