---
title: Vectorization in fitdistrplus
author: Bruce Kendall
date: '2018-11-29'
categories:
  - Code development
  - Parameter estimation
tags:
  - dispersal
slug: vectorization-in-fitdistrplus
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

I need to figure out how **fitdistr** is vectorizing the calls to the distribution functions, so I can appropriately build the 2Dt function and think sensibly about the truncated distributions. 

The gory details are below, but in summary:

- The parameter values are never vectorized
- `x` and `q` (passed to `ddist` and `pdist` respectively) may be vectors
- `x` and `q` may have length zero (which will need to be trapped for when I'm not just passing them on to a predefined distribution) (I don't know why!)
- `x` and `q` may have values NA, NaN, Inf (again, I'll need to do some trapping)

So I'll make a distribution that reports its inputs:
```{r mylnorm}
dmylnorm <- function(x, meanlog = 0, sdlog = 1, log = FALSE) {
  cat("In dmylnorm \n")
  cat("x:", x, "\n")
  cat("meanlog:", meanlog, "\n")
  cat("sdlog:", sdlog, "\n")
  dlnorm(x, meanlog, sdlog, log)
}
dmylnorm(1:10)
dmylnorm(1, (1:10)/10)

pmylnorm <- function(q, meanlog = 0, sdlog = 1, lower.tail = TRUE, log.p = FALSE) {
  cat("In pmylnorm \n")
  cat("q:", q, "\n")
  cat("meanlog:", meanlog, "\n")
  cat("sdlog:", sdlog, "\n")
  plnorm(q, meanlog, sdlog, lower.tail, log.p)
}
```

So let's generate some data and run it through `fitdist`:
```{r testfit}
mydat <- rlnorm(20) 
mydat
fitdist(mydat, "mylnorm", start = list(meanlog = 0, sdlog = 1))
```
So it is using vector values of `x` and `q`, but not of the parameters. In fact, after the initial setup, it looks like it's not using the pdist at all!

Now let's double-check the censored data fitting:
```{r testfitcens}
mydatcens <- data.frame(left = floor(mydat), right = ceiling(mydat))
mydatcens
fitdistcens(mydatcens, "mylnorm", start = list(meanlog = 0, sdlog = 1))
```
