---
title: Getting start values from moments
author: Bruce Kendall
date: '2018-11-29'
slug: getting-start-values-form-moments
categories:
  - Code development
  - Parameter estimation
tags:
  - dispersal
---

It is possible to get start values from the matching moments fit. For example:
```{r}
library(fitdistrplus)
library(actuar)
x4  <-  rpareto(1000, 6, 2)
s4 <- fitdist(x4, "pareto", "mme", order=c(1, 2), memp=function(x, order) emm(x, order))
s4$estimate
fitdist(x4, "pareto", start = as.list(s4$estimate))
```
(`emm` is in the **actuar** library; `order` needs to have a length that matches the number of parameters). But note that this requires the existence of a `mdist` function that computes the theoretical moments, given parameters. This is certainly easier than inverting the formulas for the moments from Wikipedia! **actuar** provides these for some distributions.

There doesn't seem to be a general theoretical approach to moments of truncated distributions. But since I'm only looking at lower truncation of distributions with non-negative support, I can probably get close enough by using a vector of 10 points between zero and the truncation level to get the "missing" part that needs to be removed from the theoretical moment (although there will be a bit of tricky scaling to address).

Let's double check that this works with censored data:
```{r cens}
x5 <- data.frame(left = floor(x4), right = ceiling(x4))
s5 <- try(fitdistcens(x5, "pareto", "mme", order=c(1, 2), memp=function(x, order) emm(x, order)))
# s5$estimate
# fitdist(x5, "pareto", start = as.list(s5$estimate))
```
In fact this doesn't work! `fitdistcens` only uses mle.

We could try this:
```{r cens2}
x5a <- apply(x5, 1, mean)
s5 <- try(fitdist(x5a, "pareto", "mme", order=c(1, 2), memp=function(x, order) emm(x, order)))
```
Nope, that isn't robust (in this case probably because there's so little variation in the censored data)
