---
title: Fitting the distributions
author: Bruce Kendall
date: '2019-04-17'
slug: fitting-the-distributions
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
options(keep.source = TRUE) 
ProjectTemplate::load.project() 
```

I've written the code to automatically fit all the untruncated distributions. Here it is, in `helpers.R`:
```{r}
fit_dispersal_untruncated 
```

And here are the results of fitting it to our test dataset:
```{r, warning=FALSE, message=FALSE}
temp <- filter(disperseLer, ID == "73_0")
fit_dispersal_untruncated(temp) 
```

When I first ran it, I had set the "zero" to 0, and I got an error for gengamma. So this distribution might not be too robust.


And now, fitting all the reps:
```{r fiteach, warning=FALSE, message=FALSE, cache=FALSE}
library(tidyverse)
result <- fiteach_disp_unt(disperseLer)
result
```

There they all are! The generalized gamma didn't always converge, which is the source of the warnings and the fact that it is missing from many places.

Now lets calculate the delta-AIC within each dataset:
```{r}
result <- group_by(result, ID) %>% mutate(delta_AIC = AIC - min(AIC))
print(result, n=50) 
```
And plot the results:
```{r}
ggplot(result, aes(x=delta_AIC, group = model)) + geom_histogram() + facet_wrap(~model, scales = "free") 
```

Only the generalized gamma has consistently small delta-AIC values, suggesting that I should try harder on this. Let's look at a particular case:
```{r}
temp <- filter(disperseLer, ID == "73_0")
cens_data <- cens_dispersal_data(temp, 7)
start <- start_params(cens_data, "gengamma")
start[3] <- 0
library(flexsurv)
# dgengamma <- function(x, mu, sigma, Q) {
#   print(x)
#   print(c(mu, sigma, Q))
#   (x > 0) * flexsurv::dgengamma(x, mu, sigma, Q)
# }
fit <- try(fitdistcens(cens_data, "gengamma", start = start, silent = FALSE, lower = c(-Inf, 0, -Inf), optim.method = "L-BFGS-B", control=list(trace=0)))
fit <- try(fitdist(cens_data[,2], "gengamma", start = start, silent = FALSE, lower = c(-Inf, 0, -Inf), optim.method = "L-BFGS-B", control=list(trace=0)))
summary(fit)
fit$est
fit <- try(fitdistcens(cens_data, "gengamma", start = as.list(fit$est), silent = FALSE, lower = c(-Inf, 0, -Inf), optim.method = "L-BFGS-B", control=list(trace=0)))
summary(fit)
```


Got it to fit almost everywhere, and still is always better than 4. Some cases where it is 2, suggesting a limiting distribution fits equally. But sometimes it's at zero.

```{r}
result2 <- filter(result, ID != "90_1", ID!="79_0")
plot(filter(result2, model == "gengamma")$delta_AIC, filter(result2, model == "weibull")$delta_AIC, xlab = "gengamma delta_AIC", ylab = "weibull delta_AIC")
plot(filter(result2, model == "gengamma")$delta_AIC, filter(result2, model == "gamma")$delta_AIC, xlab = "gengamma delta_AIC", ylab = "gamma delta_AIC")
plot(filter(result2, model == "gengamma")$delta_AIC, filter(result2, model == "lnorm")$delta_AIC, xlab = "gengamma delta_AIC", ylab = "lnorm delta_AIC")
```
