---
title: Distribution of Ler variability
author: Bruce Kendall
date: '2019-05-19'
slug: distribution-of-ler-variability
categories:
  - Analysis
tags:
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

So, last night's analysis suggests that there's not enough spread variability in the model. But it varied from run to run. So lets do a bunch of runs using replicate and see how far off we are.

```{r params}
n_init <- 50
Ler_params$gap_size <- 0 
controls <- list(
  n_reps = 10,
  DS_seeds = TRUE,
  ES_seeds = TRUE,
  kernel_stoch = TRUE,
  kernel_stoch_pots = TRUE,
  seed_sampling = TRUE,
  pot_width = 7
)
```

The iteration and analysis, as a function to pass to replicate:
```{r rep_func}
sim_mean_var <- function() {
  Adults <- matrix(n_init, controls$n_reps, 1)
  for (i in 1:6) {
    Adults <- iterate_genotype(Adults, Ler_params, controls)
  }
  npot <- ncol(Adults)
  rep_sum <- t(apply(Adults[, npot:1], 1, cummax))[, npot:1]
  maxd <- apply(rep_sum, 1, function(x) max((1:length(x))[x > 0]))
  maxd <- maxd[is.finite(maxd)]
  result <- c(mean(maxd), var(maxd))
  names(result) <- c("Mean", "Variance")
  result
}
```

Test the function:
```{r func_test}
sim_mean_var()
```

Do the replication
```{r replicate}
nruns <- 100
rep_spread_stats <- t(replicate(nruns, sim_mean_var(), simplify = TRUE))
rep_spread_stats
```
I had to trap for cases where the population had gone extinct by gen 6.

Make a plot.
```{r plot}
rep_spread_stats <- as.data.frame(rep_spread_stats) 
maxd_data <- pull(subset(LerC_spread, Gap == "0p" & Generation == 6), Furthest)

ggplot(rep_spread_stats, aes(x = Mean)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = mean(maxd_data))

ggplot(rep_spread_stats, aes(x = Variance)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = var(maxd_data))
```
So, the distibution of variances is really skewed and there are some really large values. but the data are right in there. Here is the data: `r c(mean(maxd_data), var(maxd_data))`. And here is the mean of the simulations: `r apply(rep_spread_stats, 2, mean)`. If I trim the most extreme values we get `r apply(rep_spread_stats, 2, mean, trim = 1/nruns)`

Just for grins, let's see what happens if we turn off kernel stochasticity.
```{r replicate2}
nruns <- 100
controls$kernel_stoch <- FALSE
rep_spread_stats <- t(replicate(nruns, sim_mean_var(), simplify = TRUE))
```
```{r plot2}
rep_spread_stats <- as.data.frame(rep_spread_stats) 
ggplot(rep_spread_stats, aes(x = Mean)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = mean(maxd_data))

ggplot(rep_spread_stats, aes(x = Variance)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = var(maxd_data))
```
That shoots it all to heck! Let's try shutting off the others one at a time.

No seed sampling:
```{r replicate3}
nruns <- 100
controls$kernel_stoch <- TRUE
controls$seed_sampling <- FALSE
rep_spread_stats <- t(replicate(nruns, sim_mean_var(), simplify = TRUE))
```
```{r plot3}
rep_spread_stats <- as.data.frame(rep_spread_stats) 
ggplot(rep_spread_stats, aes(x = Mean)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = mean(maxd_data))

ggplot(rep_spread_stats, aes(x = Variance)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = var(maxd_data))
```

No ES:
```{r replicate4}
nruns <- 100
controls$ES_seeds <- FALSE
controls$seed_sampling <- TRUE
rep_spread_stats <- t(replicate(nruns, sim_mean_var(), simplify = TRUE))
```
```{r plot4}
rep_spread_stats <- as.data.frame(rep_spread_stats) 
ggplot(rep_spread_stats, aes(x = Mean)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = mean(maxd_data))

ggplot(rep_spread_stats, aes(x = Variance)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = var(maxd_data))
```

No DS:
```{r replicate5}
nruns <- 100
controls$DS_seeds <- FALSE
controls$ES_seeds <- TRUE
rep_spread_stats <- t(replicate(nruns, sim_mean_var(), simplify = TRUE))
```
```{r plot5}
rep_spread_stats <- as.data.frame(rep_spread_stats) 
ggplot(rep_spread_stats, aes(x = Mean)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = mean(maxd_data))

ggplot(rep_spread_stats, aes(x = Variance)) + geom_histogram() + 
  geom_vline(colour = "red", xintercept = var(maxd_data))
```

Bottom line conclusions:

* Kernel stochasticity increases the mean spread rate
* Both kernel stochasticity and seed sampling greatly increases the variance in spread rate
* Both ES and DS in seed production decrease the mean spread rate, but have little impact on the variance.
