---
title: Finalizing dispersal (for now)
author: Bruce Kendall
date: '2019-04-25'
slug: finalizing-dispersal-for-now
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

Time is running short for the Davis presentation. So, at least for now, I will accept the generalized gamma distribution and the subset of data to which it fits, and move forward with creating the model.


What I need to do:

- Double check that that the replicate-specific model fits better than a single set of parameters for all the data
- Plot the resulting dispersal kernels
- Plot the patterns of covariance in parameters
- Quantify the meta-distribution of parameters, to use in the model

# Evidence of heterogeneity
The replicates to which `gengamma` didn't fit were 79_0 and 90_1. So drop these from the dataset:
```{r data_restrict}
disperseLer2 <- filter(disperseLer, ID != "79_0", ID != "90_1")
```
Now fit the combined data:
```{r fitall}
fitall <- fit_dispersal_untruncated(disperseLer2, model = "gengamma")
fitall[, -1]
```

Now fit the combined data and sum the AIC:
```{r fiteach, message=FALSE}
fiteach <- fiteach_disp_unt(disperseLer2, model = "gengamma")
sum(fiteach$AIC) 
```
The replicate-specific fits have a cumulative AIC that is massively smaller than the fit to the lumped data (by `r round(fitall$AIC - sum(fiteach$AIC))` AIC units!), so we are safe to assume heterogeneity among replicates.

# Plotting the kernels
Unfortunately, it does not appear that `stat_function()` in **ggplot2** has a straightforward way of iteratively building functions from a table of parameters. So let's try a loop:
```{r kernel_plots, warning=FALSE}
p <- ggplot(data = data.frame(x = c(0, 20)), mapping = aes(x = x)) 
for (i in 1:nrow(fiteach)) {
  pvec <- as.numeric(fiteach[i, 4:6])
  plist <- list(mu = pvec[1], sigma = pvec[2], Q= pvec[3])
  p <- p + stat_function(fun = dgengamma, args = plist)
}
p + xlab("Distance (cm)") + ylab("Probability density")
```

# Patterns in the parameters
First a scatterplot:
```{r plotmatrix}
params <- fiteach[, 4:6]
names(params) <- c("mu", "sigma", "Q")
pairs(params)
```

There's one major outlier (with a high $Q$ and low $\sigma$). But, even if I remove that point, the pattern looks rather more complex than just a multivariate normal. So I think that, rather than a parametric model for the kernel parameters, I'll just draw from the "observed" values. If I wanted to get fancy I could take into account the uncertainty associated with each estimate as well... but that's for another day (if ever).
