---
title: Drop bad RIL reps
author: Bruce Kendall
date: '2019-11-20'
slug: drop-bad-ril-reps
categories:
  - Parameter estimation
tags:
  - dispersal
  - RIL
---
```{r setup2, echo=FALSE, message=FALSE, cache=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```
First, we need to make sure that we are getting enough iterates in the fit:
```{r}
controls <- list(maxit = 1000)
```
We also clear out reps with not enough seeds (from 5/24/19), as well as dropping the reps that won't converge  on the generalized gamma (19, 40, and 77)
```{r}
n_min <- 10 # Set the minimum number of dispersing seeds
dispersing_seeds <- group_by(disperseRIL, ID) %>% 
  filter(Distance > 4) %>% 
  filter(!(ID %in% c("19", "40", "77"))) %>%
  summarize(tot_seeds=sum(Seedlings))
good_reps <- filter(dispersing_seeds, tot_seeds >= n_min) %>%
  pull(ID)
good_reps
disperseRILgood <- filter(disperseRIL, ID %in% good_reps)
```
So now fit all the reps.
```{r fitall, cache=TRUE, warning=FALSE, message=FALSE, error=FALSE}
RIL_list <- levels(disperseRILgood$RIL)
fiteach <- NULL 
for (i in RIL_list) {
  disperseRILi <- filter(disperseRILgood, RIL == i)
  fiteachi <- cbind(RIL = i, 
                    fiteach_disp_unt(disperseRILi,control=controls))
  fiteach <- rbind(fiteach, fiteachi)
} 
```

```{r}
fiteach[, 1:7]
```


Now lets calculate the delta-AIC within each dataset:
```{r deltaAIC}
result <- group_by(fiteach, ID) %>% mutate(delta_AIC = AIC - min(AIC))
ggplot(result, aes(x=delta_AIC, group = model)) + geom_histogram() + 
  facet_wrap(~model, scales = "free") 
```

Once again, we confirm that the generalized gamma is the only model that always has a low $\Delta$-AIC.

Let's look at the fitted kernels:
```{r, warning=FALSE}
fiteachgg <- filter(fiteach, model == "gengamma")
p <- ggplot(data = data.frame(x = c(0, 20)), mapping = aes(x = x)) 
for (i in 1:nrow(fiteachgg)) {
  pvec <- as.numeric(fiteachgg[i, 5:7])
  plist <- list(mu = pvec[1], sigma = pvec[2], Q= pvec[3])
  p <- p + stat_function(fun = dgengamma, args = plist, color = fiteachgg$RIL[i])
}
p + xlab("Distance (cm)") + ylab("Probability density")
```


The last question involves checking that within-genotype heterogeneity exists by comparing AIC.
First calculate summed AICs:
```{r}
AIC_sums <- group_by(fiteachgg, RIL) %>%
  summarize(AIC_sum = sum(AIC))
head(disperseRILgood)
```
And now calculate the AICs of the homogeneous models for each RIL:
```{r, cache=FALSE, eval=TRUE}
RIL_list <- levels(disperseRILgood$RIL) 
fitall <- NULL 
for (i in RIL_list) {
  disperseRILi <- filter(disperseRILgood, RIL == i)
  fitalli <- cbind(RIL = i,
                    fit_dispersal_untruncated(disperseRILi,
                                              model = "gengamma",
                                              control=controls))
  fitall <- rbind(fitall, fitalli)
}
```

And finally, the Delta-AICs:
```{r}
fitall
fitall$AIC - AIC_sums$AIC_sum
```
Positive values mean that the heterogeneous model is better. 
9 RILs show strong evidence of heterogeneity; in 3 RILs the homogeneous model is better; in 2 the two models are indistinguishable.
