---
title: Final final dispersal
author: Bruce Kendall
date: '2019-04-26'
slug: final-final-dispersal
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

A few things. 

1. On further reflection, I think that I do indeed want to use multivariate normal for generating dispersal kernels.
2. I need to find the "odd" fit and drop that from the analysis
3. For "fitall," I need to check the other distributions; maybe one of them is better.

# Eliminate odd fit
First replicate yesterday's analysis and print them out with IDs:
```{r showall}
disperseLer2 <- filter(disperseLer, ID != "79_0", ID != "90_1")
fiteach <- fiteach_disp_unt(disperseLer2, model = "gengamma")
fiteach
```
It's 131_0 that has off the wall parameter values. So let's drop that:
```{r restrict}
fiteach <- filter(fiteach, ID != "131_0")
disperseLer2 <- filter(disperseLer, ID != "79_0", ID != "90_1", ID != "131_0")
```

# Check other distributions for lumped data
See if there's another distribution that does better on the combined data:
```{r fitall, warning=FALSE}
fitall <- fit_dispersal_untruncated(disperseLer2)
fitall[, -1]
```
The generalized gamma and Weibull are essentially identical; for simplicity stick with the generalized gamma.

# MVN statistics
Add better names:
```{r rename}
names(fiteach)[4:6] <- c("mu", "sigma", "Q")
```
Calculate the mean and covariance matrix:
```{r cov}
apply(fiteach[, 4:6], 2, mean)
cov(fiteach[, 4:6])
```
It's interesting that the mean parameters are not all that different from the parameters estimated by fitting to the combined data.

# Implementation details
There are functions to implement random numbers in **MASS** and **mvtnorm**. In `gengamma` we need to ensure that $\sigma > 0$, so we'll need to trap for that and reject those instances.

The covariance matrix needs to be positive definite (symmetric, with positive eigenvalues). Let's check that:
```{r posdef}
eigen(cov(fiteach[, 4:6]))$values
```
