---
title: Beta-binomial distribution of fraction dispersing
author: Bruce Kendall
date: '2019-05-18'
slug: beta-binomial-distribution-of-fraction-dispersing
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

I never looked at how to calculate the parameters for the fraction dispersing. The trick may be the `betabinomial` function in the **VGAM** library.

The first step is to amalgamate all the dispersing seeds from `disperseLer`:
```{r munge}
nondispersers <- subset(disperseLer, Pot == 0, c("ID", "Seedlings"))
dispersers <- filter(disperseLer, Pot == 1) %>% group_by(ID) %>% summarise(dispersers = sum(Seedlings))
disperse_num <- merge(nondispersers, dispersers)
names(disperse_num)[2] <- "nondispersers"
disperse_num$dispersers <- round(2 * disperse_num$dispersers)
```

So now we make an intercept-only model with vgam.
```{r vgam}
library(VGAM)
bbfit <- vglm(cbind(dispersers, nondispersers) ~ 1, betabinomial, data = disperse_num)
summary(bbfit)
Coef(bbfit)
```
The parameterization in vgam is somewhat unconventional. mu is simply the mean of the beta distribution, alpha/(alpha + beta), but rho is 1/(1 + alpha + beta) where alpha and beta are the conventional shape parameters.

To get to the variance parameter, we need to transform this.
$$\begin{align}
V &= \frac{\alpha \beta}{(\alpha + \beta)^2 (\alpha + \beta + 1)}\\
&= \mu (1-\mu) \rho.
\end{align}$$

So we have
```{r lastbit}
fd_mean <- Coef(bbfit)[1]
fd_stdv <- sqrt(fd_mean * (1 - fd_mean) * Coef(bbfit)[2])
fd_mean
fd_stdv
```
