---
title: RIL dispersal kernels
author: Bruce Kendall
date: '2019-11-14'
slug: ril-dispersal-kernels
categories:
  - Parameter estimation
tags:
  - RIL
  - dispersal
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

I'm not entirely satisfied with the prior post on the RIL dispersal kernels.
So let me re-run some things and see what we understand.

First, we need to make sure that we are getting enough iterates in the fit:
```{r}
controls <- list(maxit = 1000)
```
We also clear out reps with not enough seeds (from 5/24/19):
```{r}
n_min <- 10 # Set the minimum number of dispersing seeds
dispersing_seeds <- group_by(disperseRIL, ID) %>% 
  filter(Distance > 4) %>% 
  summarize(tot_seeds=sum(Seedlings))
good_reps <- filter(dispersing_seeds, tot_seeds >= n_min) %>%
  pull(ID)
disperseRILgood <- filter(disperseRIL, ID %in% good_reps)
```
So now fit all the reps.
```{r fitall, cache=TRUE, warning=FALSE, message=FALSE, error=FALSE}
RIL_list <- levels(disperseRILgood$RIL)
fiteach <- NULL 
for (i in RIL_list) {
  disperseRILi <- filter(disperseRILgood, RIL == i)
  fiteachi <- cbind(RIL = i, 
                    fiteach_disp_unt(disperseRILi,control=controls))
  fiteach <- rbind(fiteach, fiteachi)
}
```

Now lets calculate the delta-AIC within each dataset:
```{r deltaAIC}
result <- group_by(fiteach, ID) %>% mutate(delta_AIC = AIC - min(AIC))
ggplot(result, aes(x=delta_AIC, group = model)) + geom_histogram() + 
  facet_wrap(~model, scales = "free") 
```

Once again, we confirm that the generalized gamma is the only model that always has a low $\Delta$-AIC.

Let's look at the fitted kernels:
```{r, warning=FALSE}
fiteachgg <- filter(fiteach, model == "gengamma")
p <- ggplot(data = data.frame(x = c(0, 20)), mapping = aes(x = x)) 
for (i in 1:nrow(fiteachgg)) {
  pvec <- as.numeric(fiteachgg[i, 5:7])
  plist <- list(mu = pvec[1], sigma = pvec[2], Q= pvec[3])
  p <- p + stat_function(fun = dgengamma, args = plist, color = fiteachgg$RIL[i])
}
p + xlab("Distance (cm)") + ylab("Probability density")
```

There are a lot of strongly squared-off distributions, which reflect datsets that have nearly uniform distance distributions.

Let's look at the scatter of parameters:
```{r pairplot, warning=FALSE}
names(fiteachgg)[5:7] <- c("mu", "sigma", "Q")
library(GGally)
ggpairs(fiteachgg, aes(color = RIL), columns = 5:7)
```

It's somewhat hard to see from this what the within-RIL patterns look like.
Make plots by RIL:
```{r pairplot2}
for (i in RIL_list) {
  fegg <- filter(fiteachgg, RIL == i)
  print(ggpairs(fegg, columns = 5:7) + ggtitle(paste("RIL", i)))
}
```

The within-RIL patterns seem reasonable, except for the fact that occasional large values of $Q$ seem like outliers.

Let's see if there's evidence that the parameters vary across RIL:
```{r}
summary(aov(cbind(mu, sigma, Q) ~ RIL, data = fiteachgg))
```
It looks like there's clear evidence for genotypic heterogeneity only in the mean dispersal distance, although there is a lot of non-normality in the other parameters.
However, I don't see an easy way to maintain the correlations when only having one of the parameters vary by RIL.
