---
title: Half-normal distribution
author: Bruce Kendall
date: '2019-04-17'
slug: half-normal-distribution
categories:
  - Parameter estimation
tags:
  - dispersal
  - Ler
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```

Last year I wrote a half-normal distribution but didn't actually test its fitness for fitting the data. So let's try it out.

Our usual data:
```{r}
temp <- filter(disperseLer, ID == "100_0")
cens_data <- cens_dispersal_data(temp, 7)
```

Let's give it a try:
```{r, cache=TRUE}
try(fitdistcens(cens_data, "hnorm"))
```

Nope, it needs a start value. Fortunately I already provided one in `start_params()`!
```{r, cache=TRUE}
try(fitdistcens(cens_data, "hnorm", start = start_params(cens_data, "hnorm")))
```
It works!

Just out of curiousity, how good is the fit?
```{r}
f1 <- fitdistcens(cens_data, "hnorm", start = start_params(cens_data, "hnorm"))
summary(f1)
try(plot(f1))
```
Oh dear, plotting requires that the quantile form of the distribution also be defined!

But we can get a basic plot:
```{r}
try(plot(f1, NPMLE.method = "Turnbull"))
```

Note that, for this dataset, the AIC is about the same as for the gamma and Weibull!
