---
title: Modeled Ler spread statistics
author: Bruce Kendall
date: '2022-11-07'
slug: modeled-ler-spread-statistics
categories: []
tags:
  - Ler
  - dispersal
---
```{r setup2, echo=FALSE, message=FALSE, cache=TRUE, warning=FALSE}
#######################################################################################
### This chunk can be deleted from any entry that lacks R code                      ###
### Set cache=FALSE during the day if you are changing the project environment;     ###
###   set it back to TRUE at the end of the day to ensure that rebuilds don't do    ###
###   unneccesary project loads.                                                    ###
#######################################################################################

# Load the project. For some reason it has to be in a separate chunck.
ProjectTemplate::load.project()
```
Here are plots that show how the model (as of 7 Nov 2022) does at predicting the mean and variance of Ler spread after 6 generations.
Each Monte Carlo replicate has simulations of 10 populations (to match the experimental design).
Within each MC replicate, the mean and variance of maximum distance across the 10 populations is calculated.
In the plots, the violin plots show the distribution of the statistics across 1000 MC replicates.
This accounts for sampling uncertainty, given only 10 replicates in the experiment.
The quartiles of the distribution are shown as horizontal lines within each distribution.
The horizontal line across the entire plot is the value from the experimental data.

```{r make_summ, cache=TRUE, message=FALSE}
Lss <- Ler_maxd_summary(Ler_sim_maxd_1000)
```
```{r make_graphs, cache=TRUE, warning=FALSE}
Ler_maxd_viz(0, Lss, "Mean")
Ler_maxd_viz(0, Lss, "Variance")
Ler_maxd_viz(1, Lss, "Mean")
Ler_maxd_viz(1, Lss, "Variance")
Ler_maxd_viz(2, Lss, "Mean")
Ler_maxd_viz(2, Lss, "Variance")
Ler_maxd_viz(3, Lss, "Mean")
Ler_maxd_viz(3, Lss, "Variance") 
```

Some observations:

1. In all but the 3-pot gaps, all versions of the model substantially overestimate the mean spread distance
2. In the 3-pot gaps, the models with kernel sampling get the mean spread distance right on; those without kernel sampling substantially underestimate the mean spread distance.
3. In the continuous runway, all versions of the model substantially underestimate the spread variance.
4. In the 1- and 2-pot gaps, the full model estimates the spread variance the best. In 1-pot gaps, all the models with both kernel sampling and seed sampling do reasonably well; the others underestimate the variance fairly substantially. In 2-pot gaps, all the models with kernel sampling do reasonably well; the others underestimate the variance to differing degrees. 
5. In the 3-pot gaps, the models with kernel sampling estimate the spread variance well; the others greatly underestimate it.
